### 2013-12-07 paper/paper.tex
moving the paper things into the paper folder

\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{wasysym}
\usepackage{algorithm}
\usepackage{algpseudocode}

\newcommand{\etal}{\emph{et al.}}

\begin{document}
\title{K Nearest Neighbour Search Using Space-filling Curves}
\author{Camara Lerner \hspace{2cm} Christopher Rabl \\
  Department of Mathematics and Computer Science\\
  University of Lethbridge, Canada}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Hilbert Curve}

The Hilbert Curve is a space filling curve that makes the next point in the curve only one unit distance from the previous point. The curve is defined based on the amount of space the curve fills. The space that the curve fills can be in any dimensionally space $(n)$ with the length of each dimension filled equal to $2^m$. This means that for each point in the curve the maximum bit size for each dimension is $(m)$~\cite{Hamilton:2006}.

The curve labels the points from $0$ to $2^{m*n} - 1$ in the order that the curve visits each point. The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner~\cite{Hamilton:2006}.

With Hilbert curves that which uses only one bit called the unit Hilbert curve, is reflected and rotated and moved around to fill the space of the larger Hilbert Curve, and then these unit Hilbert curves are connected to make the overall Hilbert curve. The only constraint on the unit Hilbert curves is that the start and end points must match up with their neighbouring unit Hilbert curves. The same idea is done with gray code, but they  can only be reflected and reversed. These facts are used to inverse the operation of and get the point from the index in the list for both gray code and Hilbert indexing. Hilbert curves also can use information from gray code in determining the changes to the unit Hilbert curve~\cite{Hamilton:2006}.

\flushleft
Formula for Gray Code is in \ref{gray-code}.
\begin{equation}
  \label{gray-code}
  gc(i) = i \oplus \left( i \gg 1 \right) 
\end{equation}

Formula for $g$ is displayed in \ref{g}.
\begin{equation}
  \label{g}
  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
\end{equation}

Formula for $d(i)$ is displayed in \ref{d}.
\begin{equation}
  \label{d}
  d(i) = 
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
      g(i) \pmod{n} & \quad \text{if $i$ is odd}
    \end{cases}
\end{equation}

Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
\begin{equation}
  \label{right-rotate}
  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
\end{equation}

Formula for $e$ is displayed in \ref{e}.
\begin{equation}
  \label{e}
  e(i) =  
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
    \end{cases}
\end{equation}

\begin{algorithm}
  \caption{The algorithm for calculating the Gray Code Inverse of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
  \label{gray-code-inverse}
  \begin{algorithmic}[1]
    \Require $g \in \mathbb{N}$
    \Return $i \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
    \State $ m \leftarrow \text{ number of bits to represent } g$ 
    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
    \While{$ j < m $ }
    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
    \State $ j \leftarrow j + 1$ 
    \EndWhile
    % \oplus is exclusive or
    % \ll left shift
    % \gg right shift
    % \mid bitwise or
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Calculating the Hilbert index given any dimensional point ${\bf p}$ 
    of size $n$ as long as the bits used within an index of the hilbert curve 
    space is specified in ($m$).}
  \label{hilbert-point-to-index}
  \begin{algorithmic}[1]
    \Require $n, m \in \mathbb{N} - \{0\} \text{ and a point } {\bf p} \in \mathbb{N}^n$
    \Return $h \in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
    \State $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $
    \For{$j\gets m - 1, 0$} 
    \State $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ 
    \State $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ 
    \State $ w \leftarrow gc^{-1} \left( l \right)$ 
    \State $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ 
    \State $ d \leftarrow d + d \left( w \right) + 1 \pmod{n}$ 
    \State $ h \leftarrow \left( h \ll n \right) | w $
    \EndFor
  \end{algorithmic}
\end{algorithm}

\bibliographystyle{amsplain}
\bibliography{mybiblio}

\end{document}